---
title: "MG_qiimeGraph_redos"
author: "Madi"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(qiime2R)
library(tidyverse)
library(cowplot)
library(magrittr)
library(vegan)
library(viridis)
```

**Functions**
```{r}
## table construction and file prep for ggplot
file_prep <- function(metadata_fp,
                      unweighted_unifrac_fp,
                      weighted_unifrac_fp,
                      shannon_fp,
                      faith_fp){
  ## metadata file
  full_metadata <- read_tsv(metadata_fp)
  full_metadata %>% 
    select(sampleid, age, cage_number, sex) %>% 
    mutate(cage_number = as.factor(cage_number)) %>% 
    mutate(age = as.factor(age)) -> metadata
  ## unweighted unifrac 
  unweighted_unifrac <- read_qza(unweighted_unifrac_fp)$data
  unweighted_unifrac_var <- unweighted_unifrac$ProportionExplained
  unweighted_unifrac <- unweighted_unifrac$Vectors
  names(unweighted_unifrac)[names(unweighted_unifrac) == 'SampleID'] <- 'sampleid'
  ## weighted unifrac
  weighted_unifrac <- read_qza(weighted_unifrac_fp)$data
  weighted_unifrac_var <- weighted_unifrac$ProportionExplained
  weighted_unifrac <- weighted_unifrac$Vectors
  names(weighted_unifrac)[names(weighted_unifrac) == 'SampleID'] <- 'sampleid'
  ## shannon
  shannon <- read_qza(shannon_fp)$data %>% 
             rownames_to_column("sampleid")
  ## faith's pd
  faith_pd <- read_qza(faith_fp)$data
  names(faith_pd)[names(faith_pd) == 'V1'] <- 'sampleid'
  names(faith_pd)[names(faith_pd) == 'V2'] <- 'faith_pd'
  ## creating tables for beta diversity ggplot construction
  ## unweighted unifrac
  unweighted_unifrac %>% 
    left_join(metadata, by = 'sampleid') %>% 
    left_join(shannon, by = 'sampleid') %>% 
    left_join(faith_pd, by = 'sampleid') -> unweighted_biom
  ## weighted unifrac
  weighted_unifrac %>% 
    left_join(metadata, by = 'sampleid') %>% 
    left_join(shannon, by = 'sampleid') %>% 
    left_join(faith_pd, by = 'sampleid') -> weighted_biom
  ## creating a list of outputs
  my_list <- list(Metadata = metadata,
                  UnweightedUniFrac = unweighted_biom,
                  UnweightedProportionExplained = unweighted_unifrac_var,
                  WeightedUniFrac = weighted_biom,
                  WeightedProportionExplained = weighted_unifrac_var)
  return(my_list)
}

## this function will pull out the percent variations from a specified column so you can add it to your pcoa plots 
pcoa_ax_lab <- function(unifrac_var, col_name){
  uni_lab <- as.character(round(unifrac_var[col_name] * 100, 2))
  uni_lab <- paste0(col_name, ' - ', uni_lab, '%')
  return(uni_lab)
}
```

**File Paths**
```{r}
metadata_FP <- '../data/misc/corrected_metadata.tsv'
unweighted_unifrac_FP <- '../data/qiime/core-metrics-corrected-MG/unweighted_unifrac_pcoa_results.qza'
weighted_unifrac_FP <- '../data/qiime/core-metrics-corrected-MG/weighted_unifrac_pcoa_results.qza'
shannon_FP <- '../data/qiime/core-metrics-corrected-MG/shannon_vector.qza'
faith_FP <- '../data/qiime/core-metrics-corrected-MG/faith_pd_vector.qza'
```

**Beta Diversity File Prep and Extraction from Named List**
```{r}
beta_files <- file_prep(metadata_FP,
                        unweighted_unifrac_FP,
                        weighted_unifrac_FP,
                        shannon_FP,
                        faith_FP)

metadata <- beta_files$Metadata
unweighted_unifrac <- beta_files$UnweightedUniFrac
uu_prop_exp <- beta_files$UnweightedProportionExplained
weighted_unifrac <- beta_files$WeightedUniFrac
wu_prop_exp <- beta_files$WeightedProportionExplained
```


**GGplot construction of my unweighted Unifrac by age**
```{r, fig.height=7, fig.width=10}
unweighted_unifrac %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = age, size = shannon_entropy, shape = sex), alpha = 0.8) +
  scale_color_viridis(option = "H", name = 'Age', discrete = TRUE) +
  theme_bw(base_size = 14) +
  scale_size_continuous(name = 'Shannon Diversity') +
  scale_shape_manual(values = c(1,16), name = 'Sex') +
  ggtitle('Unweighted Unifrac') -> unweighted_unifrac_age

unweighted_unifrac_age
```

**Reading in weighted Unifrac from Qiime2**
```{r}
weighted_unifrac <- read_qza(weighted_unifrac_fp)$data

weighted_unifrac_var <- weighted_unifrac$ProportionExplained

weighted_unifrac <- weighted_unifrac$Vectors

names(weighted_unifrac)[names(weighted_unifrac) == 'SampleID'] <- 'sampleid'
```

**Putting weighted tables together so I can graph them**
```{r}
weighted_unifrac %>% 
  left_join(metadata_red, by = 'sampleid') %>% 
  left_join(shannon, by = 'sampleid') -> weighted_unifrac
```

**GGplot construction of my weighted Unifrac by age**
```{r, fig.height=7, fig.width=10}
weighted_unifrac %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = age, shape = sex, size = shannon_entropy), alpha = 0.8) +
  scale_color_viridis(option = "H", name = 'Age', discrete = TRUE)+
  theme_bw(base_size = 14) +
  scale_size_continuous(name = 'Shannon Diversity') +
  scale_shape_manual(values = c(1,16), name = 'Sex') +
  ggtitle('Weighted Unifrac') -> weighted_unifrac_age

weighted_unifrac_age
```

**Ggplot for unweighted Unifrac by cage number**
this is too look at/hopefully eliminate cage effects in the results. 
```{r, fig.height=7, fig.width=10}
unweighted_unifrac %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = cage_number, size = shannon_entropy, shape = sex), alpha = 0.8) +
  scale_color_viridis(option = "C", name = 'Cage Number', discrete = TRUE) +
  theme_bw(base_size = 14) +
  scale_size_continuous(name = 'Shannon Diversity') +
  scale_shape_manual(values = c(1,16), name = 'Sex') +
  ggtitle('Unweighted Unifrac') -> unweighted_unifrac_cageNum

unweighted_unifrac_cageNum
```

**Ggplot for weighted Unifrac by cage number**
```{r, fig.height=7, fig.width=10}
weighted_unifrac %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = cage_number, shape = sex, size = shannon_entropy), alpha = 0.8) +
  scale_color_viridis(option = "C", name = 'Cage Number', discrete = TRUE)+
  theme_bw(base_size = 14) +
  scale_size_continuous(name = 'Shannon Diversity') +
  scale_shape_manual(values = c(1,16), name = 'Sex') +
  ggtitle('Weighted Unifrac') -> weighted_unifrac_cageNum

weighted_unifrac_cageNum
```

**Shannon Diversity**
this is an attempt at putting the shannon diversity graph together. Shannon diversity values fall between 0 and 1, with lower values meaning that there's less diversity and higher values meaning that there's more diversity in the community of microbes that is being measured. So, we'd expect to see higher shannon diversity values around day 14 than at day 7 (which we do) and that the overall trend line will be moving upwards (as it does). Shannon diversity takes into account community richness and evenness. 
```{r, fig.height=4.5, fig.width=7}
metadata %>% 
  left_join(shannon, by = 'sampleid') -> shannon_div

shannon_div %>% 
  ggplot(aes(x = age, y = shannon_entropy)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE, color = 'blue') +
  scale_x_continuous(breaks = c(7, 8, 9, 10, 11, 12, 13, 14)) +
  theme_bw(base_size = 14) +
  ggtitle('Shannon Diversity') +
  xlab('Age') +
  ylab('Shannon Diversity') -> shannon_plot

shannon_plot
```


**Faith-PD Alpha Diversity**
reading in my faith-pd data and attempting to make a boxplot out of that....I did it!! :). casey said no color so you get no color
```{r, fig.height=5, fig.width=8}
faith_pd <- read_qza(faith_fp)$data

names(faith_pd)[names(faith_pd) == 'V1'] <- 'sampleid'
names(faith_pd)[names(faith_pd) == 'V2'] <- 'faith_pd'

metadata %>% 
  left_join(faith_pd, by = 'sampleid') -> faith_pd

faith_pd %>% 
  ggplot(aes(x = age, y = faith_pd)) +
  geom_boxplot(aes(group = age), outlier.shape = NA) +
  geom_jitter(width = 0.1, height = 0) +
  scale_x_continuous(breaks = 7:14) +
  geom_smooth(se = FALSE) +
  theme_bw(base_size = 16) +
  ggtitle("Faith's Phylogenic Diversity") +
  xlab('Age') +
  ylab("Faith's PD") -> faith_plot

faith_plot
```

**Reading in weighted Unifrac from Qiime2 for Faith PD plots**
```{r}
weighted_unifrac_faith <- read_qza(weighted_unifrac_fp)$data

weighted_unifrac_faith <- weighted_unifrac_faith$Vectors

names(weighted_unifrac_faith)[names(weighted_unifrac_faith) == 'SampleID'] <- 'sampleid'

faith_pd <- read_qza(faith_fp)$data

names(faith_pd)[names(faith_pd) == 'V1'] <- 'sampleid'
names(faith_pd)[names(faith_pd) == 'V2'] <- 'faith_pd'
```


**Putting weighted tables together so I can graph them - Faith PD version**
```{r}
weighted_unifrac_faith %>% 
  left_join(metadata_red, by = 'sampleid') %>% 
  left_join(faith_pd, by = 'sampleid') -> weighted_unifrac_faith
```


**GGplot construction of my weighted Unifrac by age with Faith PD**
```{r, fig.height=7, fig.width=10}
w_uni_xlab <- pcoa_ax_lab(weighted_unifrac_var, 'PC1')
w_uni_ylab <- pcoa_ax_lab(weighted_unifrac_var, 'PC2')

weighted_unifrac_faith %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = age, size = faith_pd), alpha = 0.8) +
  scale_color_viridis(option = "H", name = 'Age', discrete = TRUE)+
  theme_bw(base_size = 16) +
  scale_size_continuous(name = "Faith's PD") +
  ggtitle('Weighted Unifrac by Age') +
  labs(x = w_uni_xlab, y = w_uni_ylab) -> wu_plot_faith_age

wu_plot_faith_age
```


**Ggplot construction of my weighted Unifrac by cage number with Faith PD**
```{r, fig.height=7, fig.width=10}
weighted_unifrac_faith %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = cage_number, size = faith_pd), alpha = 0.8) +
  scale_color_viridis(option = "C", name = 'Cage Number', discrete = TRUE) +
  theme_bw(base_size = 16) +
  scale_size_continuous(name = "Faith's PD") +
  ggtitle('Weighted Unifrac by Cage Number') +
  labs(x = w_uni_xlab, y = w_uni_ylab) -> wu_plot_faith_cage

wu_plot_faith_cage
```

**Saving Plot Outputs**
```{r}
ggsave('unweighted_unifrac_age.pdf',
       plot = unweighted_unifrac_age, 
       width = 10,
       height = 7,
       path = '../plots')
ggsave('weighted_unifrac_age.pdf',
       plot = weighted_unifrac_age, 
       width = 10,
       height = 7,
       path = '../plots')
ggsave('unweighted_unifrac_cage.pdf',
       plot = unweighted_unifrac_cageNum, 
       width = 10,
       height = 7,
       path = '../plots')
ggsave('weighted_unifrac_cage.pdf',
       plot = weighted_unifrac_cageNum, 
       width = 10,
       height = 7,
       path = '../plots')
ggsave('shannon_entropy.pdf',
       plot = shannon_plot, 
       width = 7,
       height = 4.5,
       path = '../plots')
ggsave('faith_pd.pdf',
       plot = faith_plot, 
       width = 8,
       height = 5,
       path = '../plots')
```

