---
title: "MG-taxabarplot"
author: "Madi"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(qiime2R)
library(tidyverse)
library(cowplot)
library(magrittr)
library(viridis)
library(microshades)
library(phyloseq)
```

**Functions**
```{r}
metadata_fixer <- function(metadata_fp) {
  # we want a function that will get rid of the second
  # line (qiime2 stuff that we don't need)
  # to do this, we need to 
  # only read in first two lines to save time.
  tmpMeta <- read_csv(metadata_fp, n_max = 2)
  mycols <- colnames(tmpMeta)
  
  #reading in metadata file, specifying that we want first two rows skipped
  ## then we're saying that we want 'mycols' to be saved as the column names 
  ## we took 'mycols' from the original metadata file (i.e. one of the rows that
  ## we skipped)
  metadata <- read_csv(metadata_fp, skip = 2, col_names = mycols)
  ## return allows you to return an object as a variable instead of just printing
  ## it 
  return(metadata)
} 
```

**Reading in needed files**
needed files are my metadata file, the qiime2 taxonomy file, and my tax-filtered table
```{r}
svs <- read_qza(file='~/neonatal-gut-microbiome/tax-filtered-MG.qza')$data

read_qza(file='~/neonatal-gut-microbiome/taxonomy-MG.qza')$data %>% 
  parse_taxonomy() -> taxonomy_MG

taxasums <- summarize_taxa(svs,taxonomy_MG)$Genus

metadata <- metadata_fixer(metadata_fp = "MG_metadata.csv")
```

**Creating my long format biom table**
needed to go from wide to long format or else it breaks ggplot
```{r}
taxasums %>% 
  as_tibble(rownames = 'taxonomy') %>% 
  gather(-taxonomy, key = sampleid, value = abund) %>% 
  left_join(metadata) %>% 
  group_by(sampleid) %>% 
  mutate(relabund=abund/sum(abund)) %>% 
  separate(taxonomy, into = c('Kingdom', 'Phylum', 'Class', 
                              'Order', 'Family', 'Genus', 
                              'Species'), sep = ' ') -> biom_long
```

**Creating my prevalence table**
```{r}
biom_long %>% 
  select(Family, relabund, sampleid) %>% 
  group_by(Family) %>% 
  summarize(prevalence = sum(relabund > 0)/nrow(metadata)) %>% 
  arrange(desc(prevalence)) -> prevalence

PREVALENCE_THRESH <- 0.7
```

**Age Taxabarplot construction**
this barplot is sectioned out by mouse age for visibility
```{r, fig.width=25, fig.height=18}
biom_long %>%
  left_join(prevalence) %>% 
  group_by(sampleid, Family, prevalence, age) %>% 
  summarize(relabund = sum(relabund),
            prevalence = sum(prevalence) > 0) %>% 
  filter(prevalence > PREVALENCE_THRESH) %>% 
  filter(Family != 'NA;') %>% 
  filter(Family != 'Mammalia;') %>% 
  filter(Family != 'uncultured;') -> family_taxa

family_taxa %>% 
  ggplot(aes(x = sampleid, y = relabund)) +
    geom_bar(aes(fill = Family), color = 'black',
             stat='identity', position='stack') +
    scale_fill_viridis(option = 'B', name = 'Family', discrete = TRUE) +
    facet_wrap(~age, scales = 'free_x') +
    theme_bw(base_size = 16) +
    xlab('Sample ID') +
    ylab('Relative Abundance')
```
**Full Taxabarplot Construction**
this taxabarplot has everything together!
```{r, fig.width=15, fig.height=15}
biom_long %>%
  left_join(prevalence) %>% 
  group_by(sampleid, Order, prevalence, age) %>% 
  summarize(relabund = sum(relabund),
            prevalence = sum(prevalence) > 0) %>% 
  filter(prevalence > PREVALENCE_THRESH) %>% 
  filter(Order != 'NA;') %>% 
  ggplot(aes(x = age, y = relabund)) +
    geom_bar(aes(fill = Order), color = 'black',
             stat='identity', position='stack') +
    scale_fill_viridis(option = "H", discrete = TRUE) +
    xlab('Mouse Age') +
    ylab('Relative Abundance')
```
**Attempting to figure out microshades for my taxa barplot**
```{r}

taxa_1 <- read_qza(file='~/neonatal-gut-microbiome/taxonomy-MG.qza')$data %>% 
  parse_taxonomy()

##taxa_1 %>% 
 ## tax_glom('Family') %>% 
 ## phyloseq::transform_sample_counts(function(x) {
 ##   x/sum(x)
 ## }) %>% 
  ##psmelt() %>% 
  ##filter(relabund > 0)

##micro_color_gen <- create_color_dfs(family_taxa, group_level = 'Phylum', subgroup_level = 'Family', cvd = TRUE)

```

